Functional vs. Object Oriented Programming

### 2015-10-10 article.tex
Tables of results and references added

The experimentation consist in static code analysis on a set of eight algorithms, namely:

\begin{itemize}
\item FASTA
\item k-nucleotide
\item Mandelbrot
\item n-body
\item pidigits
\item regex-dna
\item reverse-complement
\item spectral-norm
\end{itemize}

@@

@@

What we evince from the gathered data is that, normally, there is no big difference between the LOC for FP and OOP. Nevertheless, in certain algorithms the LOC in FP implementations are less than the half of OOP implementation.

@@

>>>>

The experimentation consist in static code analysis on a set of eight algorithms\cite{algorithms}, namely:

\begin{itemize}
\item FASTA
\item k-nucleotide
\item Mandelbrot
\item n-body
\item pi-digits
\item regex-dna
\item reverse-complement
\item spectral-norm
\end{itemize}

@@

\begin{table}[htbp]
\centering
\caption{LOC for different implementations.}
\label{tab:all}
    \begin{tabular}{rcccccccccc}

 & Haskell & Erlang & Scala & Clojure & F\#  & Python & C++ & C\#  & Go  & Java \\ 
\hline
        FASTA                & 111     & 61     & 91    & 129     & 69  & 84     & 132 & 144 & 162 & 292  \\ 
        k-nucleotide         & 159     & 67     & 194   & 266     & 104 & 156    & 117 & 168 & 175 & 143  \\ 
        Mandelbrot           & 49      & 42     & 59    & 110     & 54  & 59     & 79  & 63  & 103 & 55   \\ 
        n-body               & 166     & 114    & 117   & 189     & 98  & 89     & 132 & 100 & 127 & 138  \\ 
        Pi-digit             & 13      & 33     & 88    & 266     & 89  & 33     & 72  & 113 & 85  & 110  \\ 
        regex-dna            & 106     & 95     & 148   & 51      & 52  & 38     & 62  & 38  & 77  & 148  \\ 
        reverse-compl.   & 68      & 136    & 66    & 64      & -   & 61     & 222 & 85  & 174 & 182  \\ 
        spectral-norm        & 22      & 35     & 50    & 94      & 59  & 39     & 161 & 109 & 72  & 98  \\
\hline
    \end{tabular}
\end{table}

\begin{table}[htbp]
\centering
\caption{LOC for the two paradigms.}
\label{tab:all}
    \begin{tabular}{rcccccccccc}

 & FP & OOP \\ 
\hline
        FASTA  & 91 & 183 \\ 
        k-nucleotide & 158 & 151\\ 
        Mandelbrot & 63 & 75 \\ 
        n-body & 129 & 125 \\ 
        Pi-digit & 87 & 95 \\
        regex-dna & 82 & 65 \\ 
        reverse-compl. & 66 & 166 \\ 
        spectral-norm & 50 & 110 \\
\hline
    \end{tabular}
\end{table}

@@

What we evince from the gathered data is that, normally, there is no big difference between the LOC for FP and OOP. Nevertheless, in algorithms, like FASTA, revers-complement and spectral-norm, the LOC in FP implementations are less than the half of OOP implementation. This find explanation in \cite{fpmatter} where the benefits of FP are discussed. Among the most significative there are: the syntax (which is simpler), higher order functions and the standard list-manipulation primitives.

@@

Static code analysis on the lines of code itself is not enough to have a global vision about the two paradigms. In fact, there are many additional factors that affect productivity such as the documentation, the comments, the development time, the application domain, the experience of the developer, the time to implement modifications\cite{harrison, Haskell-vs-ada}.

For example, in Fig. \ref{fig:hudak} we notice that a very concise implementation may require a bigger documentation, like in the case of (1). Furthermore, we observe that in (1), (2), (3), (10) the development time is proportional to the LOC and this fact support our hypothesis.

It has not been possible to succeed in the experimentation concerning the CR as we could not find a suitable case study\cite{play, flink, rx}. For instance, many implementations are based on the JVM so that the developers reuse their Java implementation building an API layer for functional languages, like Scala and Clojure.

\begin{figure}[h]
\includegraphics[width=\textwidth]{4}
\caption{The result of Hudak and Jones experimentation in\cite{Haskell-vs-ada}.}
\label{fig:hudak}
\end{figure}
